SNP Analysis

### 2018-09-05 snp/SNP_Prediction_Analysis.Rmd
add Seattle SNP analysis scripts

---
title: "SNP Analysis"
author: "Sarah Killcoyne"
date: "6/13/2018"
output: html_document
---

@@

suppressPackageStartupMessages(source('lib/data_func.R'))
suppressPackageStartupMessages(source('lib/common_plots.R'))

chr.info = get.chr.lengths(file='~/tmp/hg19_info.txt')

@@

qcdata = base::merge(qcdata, sample.info[,c('Samplename','Status','Timepoint','SCA.Ratio')], by='Samplename')

ggplot(qcdata, aes(round(Ploidy,2))) + geom_histogram() + 
  facet_wrap(~SampleType, scales='free_y', ncol=1) + labs(x='Ploidy', title='Ploidy, normal vs BE')

ggplot(qcdata, aes(Purity, Goodnessoffit, color=Status)) + geom_point() + facet_wrap(~SampleType, ncol=1) + 
  labs(title='Purity~fit, normal vs BE')

ggplot(subset(qcdata, SampleType == 'BE'), aes(round(Ploidy,2), Purity, color=Status)) + geom_point() + facet_grid(~Status) + 
  labs(title='BE only, purity vs ploidy')

ggplot(qcdata, aes(SCA.Ratio, color=Status, fill=Status)) + geom_density(alpha=0.5) + labs(x='SCA Ratio', title='SCA Ratio, NP vs P')

@@

#copySegs = as.data.frame(normalize.quantiles.use.target(mergedSegs, as.vector(raw.segs)))
grid.arrange(
  ggplot( melt(raw.segs), aes(sample=value)) + stat_qq() + labs(title='Normal Q-Q plot, sWGS bins (no arms)'),
  ggplot( melt(mergedSegs), aes(sample=value)) + stat_qq() + labs(title='Normal Q-Q plot, SNP bins (no arms)'),
  ggplot( melt(copySegs), aes(sample=value)) + stat_qq() + labs(title='Normal Q-Q plot, SNP bins q-normed')
)

@@

@@

qcdata.be = subset(qcdata, SampleType == 'BE')
get.qcinfo<-function(vec) {
  endo = vec[2]
  pt = vec[1]
  y = subset(qcdata.be, PatientID == pt & EndoID == endo)
  y = y %>% group_by(Status, Timepoint) %>%  dplyr::summarise( mean(Ploidy), mean(Purity), mean(Goodnessoffit), mean(SCA.Ratio) )
  cbind.data.frame('PatientID'=pt, 'EndoID'=endo, y, stringsAsFactors = F)
}

@@

ggplot(pred.qc, aes(`mean(SCA.Ratio)`, Prob, color=Status)) + geom_point()
ggplot(pred.qc, aes(`mean(Ploidy)`, Prob, color=Status)) + geom_point()

pander(cor.test(pred.qc$`mean(SCA.Ratio)`, pred.qc$Prob))

prog = subset(pred.qc, Status == 'P')
p1 = ggplot(prog, aes(Prob)) + facet_wrap(~Timepoint, scales='free_y',nrow=1) + 
  geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
  scale_fill_gradientn(colors = myPal,  name='') + 
  labs(title='Predictions on progressors', y='n Samples', x='Unadjusted Probability')

@@

pred.qc = pred.qc %>% group_by(Status, PatientID) %>% dplyr::mutate(
  exclude=(Status == 'NP' & (length(which(Pathology == 'HGD')) > 0)),
  wgd = (Status == 'NP' & `mean(Ploidy)` > 2.7 )
)
pred.qc = pred.qc %>% group_by(Status, Samplename) %>% dplyr::mutate(
  lowsca = (Status == 'P' & (`mean(SCA.Ratio)` < 0.03 | `mean(Ploidy)` < 1.5))
)

subset(pred.qc, `mean(Ploidy)` < 1.99)

subset(pred.qc, lowsca)

@@

pred.qc.ex$quants = with(pred.qc.ex, cut(Prob, breaks=cuts))
qt = as.data.frame.matrix(table(pred.qc.ex$quants, pred.qc.ex$Status))
qt$quant = rownames(qt)
qt$P = round(qt$P/sum(qt$P),2)
qt$NP = round(qt$NP/sum(qt$NP), 2)
pander(qt)

@@

per_pt = 
  subset(pred.qc, !exclude) %>% group_by(PatientID, Status) %>% dplyr::summarise( max.Prob = max(Prob), max.RR = max(RR), max.Risk = max(Risk) )
rocpp = pROC::roc(Status ~ max.Prob, data=per_pt, auc=T, ci=T, of='thresholds')
roc.plot(rocpp) + labs(title='Max per patient - excluded HGD NP')

@@

# Exclude the low SCA patients
rocpp = pROC::roc(Status ~ Prob, data=subset(pred.qc, !lowsca), auc=T, ci=T, of='thresholds')
roc.plot(rocpp) + labs(title='Both endos - excluded low SCA P')
pROC::coords(rocpp, 'best')

# Exclude the tetraploid NP
rocpp = pROC::roc(Status ~ Prob, data=subset(pred.qc, !wgd), auc=T, ci=T, of='thresholds')
roc.plot(rocpp) + labs(title='Both endos - excluded WGD NP')
pROC::coords(rocpp, 'best')

@@

```{r regression, echo=F, message=F, eval=F}
## ----- All ----- ##
cache.dir = '~/Data/Ellie/Analysis/SNP'
file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
if (file.exists(file)) {
 message(paste("loading", file))
 load(file, verbose=T)
 model.performance(performance.at.1se, coefs)$plot + labs(title='SNP CV models')
} else {
  warning("Run 'snp-regression.R' first")
}
#save(arrayDf,labels,patient.info, file=paste(cache.dir,'snp-models.Rdata',sep='/'))

@@

```{r loo, eval=F}

slabels$Prediction = NA
slabels$RR = NA

@@

slabels
snp.roc = pROC::roc(Status~Prediction,slabels, auc=T, ci=T, of='thresholds')
roc.plot(snp.roc, title="SNP LOO Model ROC")

>>>>

---
title: "SNP Analysis"
author: "Sarah Killcoyne"
date: "6/13/2018"
output: 
  html_document: 
    toc: yes
---

@@

source('~/workspace/shallowwgs_pipeline/lib/data_func.R')
source('~/workspace/shallowwgs_pipeline/lib/common_plots.R')

chr.info = get.chr.lengths(file='/tmp/hg19_info.txt')

@@

qcdata = base::merge(qcdata, sample.info[,c('PatientID','Samplename','Status','Timepoint','SCA.Ratio')], by='Samplename')

ggplot(subset(qcdata,SampleType == 'BE'), aes(Timepoint, SCA.Ratio, color=PatientID, group=Timepoint)) + geom_boxplot() + 
  facet_wrap(~Status, scales='free_x', ncol=2) + labs(title='SCA Ratio by Timepoint') + theme(legend.position = 'none')

ggplot(subset(qcdata,SampleType == 'BE'), aes(Timepoint, SCA.Ratio, color=PatientID, group=PatientID)) + geom_point() + geom_line() + 
  facet_wrap(~Status, scales='free_x', ncol=2) + labs(title='SCA Ratio by Timepoint') + theme(legend.position = 'none')

ggplot(qcdata, aes(round(Ploidy,2))) + geom_histogram() + 
  facet_wrap(~SampleType, scales='free', ncol=1) + labs(x='Ploidy', title='Ploidy, normal vs BE')

ggplot(qcdata, aes(round(Purity,2))) + geom_histogram() + 
  facet_wrap(~SampleType, scales='free', ncol=1) + labs(x='Purity', title='Purity, normal vs BE')

ggplot(qcdata, aes(round(Ploidy,2), Purity, group=round(Ploidy,2), color=Status)) + geom_point() + 
  facet_wrap(~Status, scales='free', ncol=2) + labs(x='Ploidy', title='Ploidy vs Purity, normal vs BE')

#x=median(qcdata$SCA.Ratio,na.rm=T)+sd(qcdata$SCA.Ratio,na.rm=T)
x=0.01
y=median(qcdata$Purity,na.rm=T)-sd(qcdata$Purity,na.rm=T)/2

ggplot(qcdata, aes(Purity, SCA.Ratio, color=Status)) + geom_point() + 
         facet_wrap(~Status, ncol=1) + labs(title='Purity~SCA.Ratio, normal vs BE') + geom_vline(xintercept=y, color='grey')

ggplot(qcdata, aes(SCA.Ratio, Purity, color=Status)) + geom_point() + 
  facet_wrap(~Status) + labs(x='SCA', title='SCA vs Purity') + 
  geom_vline(xintercept = x, color='grey') + geom_hline(yintercept = y, color='grey')

table(subset(qcdata, Purity > y & SCA.Ratio < x & SampleType == 'BE')$Status)

ggplot(qcdata, aes(Purity)) + geom_histogram() + geom_vline(xintercept = y, col='red') + 
  facet_wrap(~SampleType, scales = 'free') + labs(title='Purity by sampletype')

ggplot(qcdata, aes(Purity, Goodnessoffit, color=Status)) + geom_point() + 
         facet_wrap(~SampleType, ncol=1) + labs(title='Purity~fit, normal vs BE') + geom_vline(xintercept=y, color='grey')

ggplot(subset(qcdata, SampleType == 'BE'), aes(round(Ploidy,2), Purity, color=Status)) + geom_point() + facet_grid(~Status) + 
  labs(title='BE only, purity vs ploidy') + geom_hline(yintercept=y, color='grey')

ggplot(qcdata, aes(SCA.Ratio, color=Status, fill=Status)) + facet_wrap(~Status, scales='free_y',nrow=1) +  geom_density(alpha=0.5) + labs(x='SCA Ratio', title='SCA Ratio, NP vs P')

summary(qcdata$SCA.Ratio)

@@

grid.arrange(
  ggplot( melt(raw.segs), aes(sample=value)) + stat_qq() + labs(title='Normal Q-Q plot, sWGS bins (no arms)'),
  ggplot( melt(mergedSegs), aes(sample=value)) + stat_qq() + labs(title='Normal Q-Q plot, SNP bins (no arms)'),
  ggplot( melt(copySegs), aes(sample=value)) + stat_qq() + labs(title='Normal Q-Q plot, SNP bins q-normed')
)

@@

In the gastric/blood normals `r signif(length(which(preds.normal$Prob < 0.5))/nrow(preds.normal),2)` are predicted to have <50% probability (TPR).

@@

```{r, warning=F, fig.height=8}
qcdata.be = subset(qcdata, SampleType == 'BE')
get.qcinfo<-function(vec) {
  endo = vec[2]
  pt = vec[1]
  y = subset(qcdata.be, PatientID == pt & EndoID == endo)
  y = y %>% group_by(Status, Timepoint) %>%  dplyr::summarise( mean(Ploidy), mean(Purity), mean(Goodnessoffit), mean(SCA.Ratio) )
  cbind.data.frame('PatientID'=pt, 'EndoID'=endo, y, stringsAsFactors = F)
}

@@

grid.arrange(
  ggplot(pred.qc, aes(`mean(SCA.Ratio)`, Prob, color=Status)) + geom_point(),
  ggplot(pred.qc, aes(`mean(Ploidy)`, Prob, color=Status)) + geom_point(),
ncol=2)

pander(cor.test(pred.qc$`mean(SCA.Ratio)`, pred.qc$Prob), caption='Probability shows a 0.3 correlation with % of the genome altered')
```

```{r, warning=F, fig.height=8}
prog = subset(pred.qc, Status == 'P')
p1 = ggplot(prog, aes(Prob)) + facet_wrap(~Timepoint, scales='free_y',nrow=1) + 
  geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
  scale_fill_gradientn(colors = myPal,  name='') + 
  labs(title='Predictions on progressors', y='n Samples', x='Unadjusted Probability')

@@

pred.qc = pred.qc %>% group_by(Status, PatientID) %>% dplyr::mutate(
  exclude=(Status == 'NP' & (length(which(Pathology == 'HGD')) > 0)),
  wgd = (Status == 'NP' & `mean(Ploidy)` > 2.7 )
)

pred.qc = pred.qc %>% group_by(Status, Samplename) %>% dplyr::mutate(
  lowsca = ( Status == 'P' & (`mean(SCA.Ratio)` < 0.03 & `mean(Purity)` >= 0.95))
)

# pred.qc = pred.qc %>% group_by(Status, Samplename) %>% dplyr::mutate(
#   lowsca = ( `mean(SCA.Ratio)` < 0.01 & `mean(Purity)` >= 0.95)
# )

#table(subset(pred.qc, !lowsca)$Status)/table(pred.qc$Status)

ggplot(pred.qc, aes(`mean(SCA.Ratio)`, `mean(Purity)`, color=Status)) + geom_point() + facet_grid(~Status) + 
  geom_vline(xintercept=0.01) + geom_hline(yintercept = 0.95)

ggplot(pred.qc, aes(`mean(SCA.Ratio)`))  + geom_histogram() + facet_wrap(~Status, scales='free_y')

ggplot(pred.qc, aes(`mean(Purity)`))  + geom_histogram() + facet_wrap(~Status, scales='free_y')

ggplot(pred.qc, aes(Prob)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
  scale_fill_gradientn(colors = myPal,  name='') + 
  labs(title='Predictions, all samples model', y='n Samples', x='Unadjusted Probability')

@@

pred.qc.ex$quants = with(pred.qc.ex, cut(Prob, breaks=cuts))
qt = as.data.frame.matrix(table(pred.qc.ex$quants, pred.qc.ex$Status))
qt$quant = rownames(qt)
qt$P = round(qt$P/sum(qt$P),2)
qt$NP = round(qt$NP/sum(qt$NP), 2)
pander(qt, caption='Excluding HGD NP, WGD NP, low SCA P')

@@

# per_pt = 
#   subset(pred.qc, !exclude) %>% group_by(PatientID, Status) %>% dplyr::summarise( max.Prob = max(Prob), max.RR = max(RR), max.Risk = max(Risk) )
# rocpp = pROC::roc(Status ~ max.Prob, data=per_pt, auc=T, ci=T, of='thresholds')
# roc.plot(rocpp) + labs(title='Max per patient - excluded HGD NP')

# Exclude the WGD NP
rocpp = pROC::roc(Status ~ Prob, data=subset(pred.qc, !wgd), auc=T, ci=T, of='thresholds')
roc.plot(rocpp) + labs(title='Both endos - excluded WGD NP')
pROC::coords(rocpp, 'best')

@@

# Exclude the low SCA patients
rocpp = pROC::roc(Status ~ Prob, data=subset(pred.qc, !lowsca), auc=T, ci=T, of='thresholds')
roc.plot(rocpp) + labs(title='Both endos - excluded low SCA')
pROC::coords(rocpp, 'best')

@@

```{r regression, echo=F, message=F, eval=T}
folds = 10; splits = 5
## ----- All ----- ##
cache.dir = '~/Data/Ellie/Analysis/SNP'
file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
if (file.exists(file)) {
 message(paste("loading", file))
 load(file, verbose=T)
 model.performance(performance.at.1se, coefs)$plot + labs(title='SNP CV models')
} else {
  warning("Run 'snp-regression.R' first")
}
#save(arrayDf,labels,patient.info, file=paste(cache.dir,'snp-models.Rdata',sep='/'))

@@

```{r loo, eval=T}

@@

snp.roc = pROC::roc(Status~Prediction,slabels, auc=T, ci=T, of='thresholds')
roc.plot(snp.roc, title="SNP LOO Model ROC")
